Effects of Unemployment Benefits and Uncertainty in Heterogeneous Models

### 2017-01-21 presentation/presentation.tex
Added slides with equations from the Aiyagari Model.
First draft of the presentation, plots are still missing.

	\item {
	Why the Krussell Smith model? 
	}

@@

	\item {
	Consumers 
	}

@@

@@

\section{Aggregate Uncertainty}
\subsection{}

\begin{frame}{Idiosyncratic Uncertainty}
	\begin{itemize}

	\item {
	Why the Aiyagari Model? - Explaining the 				differences from the Representative Agent Model.
	}
	\item {
	Model Mechanics - How does a change in 					idiosyncratic uncertainty affect the agents' behavior?
	}
	\item {
	What conclusions can we draw from such an analysis?
	}
	\item {
	Where are the model's limitations?
	}		
	\end{itemize}
	  \centering{\includegraphics[scale=0.30]{eq18}}

@@

>>>>

	\item {
	Why the Krussell Smith Model? 
	}

@@

\begin{frame}{Model Equations}
	\begin{itemize}

	\item {
	Agents maximize utility:
	}

\[ 
max_{\{{ c_{i,t}, k_{i,t+1} }\}_{t = 0}^{\infty}} {\mathbb{E}} \sum_{i=0}^{\infty} \beta^{t}  u(c_{i,t}) 
\]
 \[ 
 c_{i,t} + k_{i,t+1} = (1 + r_{t} - \delta) k_{i,t} + w_{t} (1 - \tau_{t})  e_{i,t} + \mu w_{t} (1 - e_{i,t})
 \]
  \[
  k_{i,t + 1} \geq \bar{k}
	\]

@@

\begin{frame}{Model Equations}
	\begin{itemize}

	\item {
	Representative Firm maximizes profit:
	}
\[ 
\underset{K_{t},L_{t}}{max} F_{t}(K_{t},L_{t}) -r_{t}K_{t} - w_{t}L_{t}
\]

	\item {
	Government redistributes taxes from employed to unemployed in terms of benefits:
	}

\[ 
\tau_{t}L_{t} = \mu(1-L_{t})
\]

\begin{frame}{System of Equations}

Euler equation:
\[ 
c_{i,t}^{- \sigma} \geq \beta (1+r -\delta){\mathbb{E}[c_{i,t+1}^{- \sigma}}]
\]
Budget constraint:
 \[ 
 c_{i,t} + k_{i,t+1} = (1 + r_{t} - \delta) k_{i,t} + w (1 - \tau)  e_{i,t} + \mu w (1 - e_{i,t})
 \]
Borrowing constraint:
  \[
  k_{i,t + 1} \geq k_{min}
	\]
Factor prices:
\[
r = \alpha z (\tfrac{L}{K})^{1-\alpha} \ \ \ w = (1-\alpha)z (\tfrac{K}{L})^{\alpha}
\]
Employment and Tax rate: 
\[
L = \tfrac{\pi_{UE}}{\pi_{UE}+\pi_{EU}} \ \ \ \ \  \tau = \mu \tfrac{1-L}{L}
\]

@@

\begin{frame}{Changes in idiosyncratic risk}
	\begin{itemize}

	\item {
	We are interested in studying the mechanics behind changes in precautionary savings. 
	}
	\item {
	We thus look at different steady states corresponding to different values of $\mu$ and study the changes of the model variables.
	}

\begin{frame}{Changes in idiosyncratic risk}
	\begin{itemize}

	\item {
	First, we look at the changes in aggregates and compare them to the representative agent case.
	}

\begin{frame}{Changes in idiosyncratic risk}

graph with log-deviations and aggregates 
  \centering{\includegraphics[scale=0.30]{eq18}}

@@

\begin{frame}{Changes in idiosyncratic risk}
	\begin{itemize}

	\item {
	As the unemployment insurance increases and idiosyncratic risk decreases, the aggregates decrease and approach the values of the representative agent model. 
	}

	\item {
Hence, with an increasing $\mu$ savings tend to decrease on average.
	}

	\item {
Looking at the distribution of capital holdings gives us an even better picture.
	}

\begin{frame}{Changes in idiosyncratic risk}

distributional plot 
  \centering{\includegraphics[scale=0.30]{eq18}}

\begin{frame}{Changes in idiosyncratic risk}

\begin{itemize} 
	\item {
We saw that the policy change has a different effect on the unemployed and the employed. 
}
	\item {
Hence, we will now look at the different variables, to get an even better image.
}
\end{itemize}

\begin{frame}{Changes in idiosyncratic risk}

big subplot with all the variables
  \centering{\includegraphics[scale=0.30]{eq18}}

\begin{frame}{Changes in idiosyncratic risk}

We can distinguish between different effects:

	\begin{itemize}
	\item {
The Income-effect: On aggregate consumers will have more income as unemployed receive additional benefits

	}	
	\item {
Tax-effect: The working are taxed to finance the unemployment benefits and to keep the government buget in balance, reducing their net-income.

	}	
	\item {
Price effect: A reduction of individual risk by a more generous UI benefit decreases precautionary savings, and therefore reduces the aggregate capital stock. This will increase the interest rate and reduce the wage rate.
	}

\begin{frame}{Changes in idiosyncratic risk}

What conclusions can we draw from this?

	\begin{itemize}
	\item {
There are diverse effects, affecting the agents' welfare in contradictory ways. Hence, we do need a different metric, in order to make qualified statements about the policy change. 
	}	
	\item {
The model neglegcts certain aspects: There is no voluntary unemployment. A change in the UI does not affect the overall unemployment rate, which contradicts empirical evidence. 
	}	
	\item {
Finally, it is important to note, that we only compare steady states. However, every policy change includes a transition phase, which might aversely affect the agents' welfare. 
	}

\section{Welfare effects}
\subsection{}

\section{Aggregate risk}
\subsection{}

\section{Conclusions}
\subsection{}
